Constraining the High Density Nuclear Symmetry Energy with Pions

### 2020-07-04 chapters/ch_data_analysis_2.tex
Something broken. Saftey backup
backup
Halfway through data analysis 1 backup
Second draft. Intro needs work

\section{Radio Isotope Beam Factory (RIBF) Facility }
\label{sec:beam}
%Cyclotron facility overview.
%Samurai line overview.
%Beam line element overview.
%Big rips beam PID. reference 
The primary and secondary beams were produced at the Radioactive Isotope Beam Factory (RIFB) facility at RIKEN, in Wako-shi, Japan. The RIBF facility starts with two primary beam types, ${}^{132}$Xe and ${}^{238}$U, which are produced by an ion-source and accelerated to progressively higher kinetic energies by 1 linear accelerator (RILAC), and 4 different cyclotrons (RRC, fRC, IRC, and SRC), until they reach a beam energy of \SI{345}{\MeVA}. Figure~\ref{fig:samuraiBeamLine} shows the later stages of the cyclotrons and the following beam lines they feed into.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{beamPID.png}
\label{fig:beampid}
\caption{Overall beam PID for all the systems. Several contaminants other than the desired secondary beam can be seen.}
\end{figure}

@@

%Figures of beam contaminants in our beam PID line. 
%Table of beam purity reference to Jon's thesis paper
The secondary beam is produced through the projectile fragmentation of the primary beam off of a \SI{3}{\milli\metre} thick, rotating Be target \cite{inflightsep}. The resulting fragments are filtered in-flight to the desired seconary beam. The in-flight separation is handled by the BigRIPS fragment separator which is shown in Fig.~\ref{fig:samuraiBeamLine}. The dipole magnets D1 and D2 act as a velocity filter, selecting on certain magnetic rigidities $\beta\rho$. Several sets of slits further purify the secondary beam quality by throwing away particles which do not focus on the right focal planes. These are the areas where the particles with different velocities focus to different locations in space, which occur at F3,F5, and F7 positions.  Each beam is tracked with the remaining part of the BigRIPS spectrometer tracking system. The particle identification of each beam is achived by the TOF-B$\rho$-$\Delta$E method described in \cite{bigrips}, where the Time of Flight (TOF) information is given by the time it takes to cross two plastic scintilators at F3 and F7 focal planes, and the $\Delta$E information is given by the MUlti-Sampling Ionization Chamber (MUSIC) \cite{music}. From this method the atomic charge, Z, and mass to charge ratio, A/Q, of each particle was measured and separate species represent two-dimensional Gaussians in this space.

Figure~\ref{fig:beampid} shows the beam PID for all the systems for events which satisfied the trigger, several contaminants other than the desired secondary beams still passed through the BigRIPS spectrometer and made it into the TPC. The beam purity of each desired secondary beam is listed in Table~\ref{tb:beams}. The desired secondary beam of interest can be selected by using an appropriate gate around the corresponding group, since the BigRIPS PID resolution is good enough to separate separate beam types. Each particle gate is selected by fitting a multivariate normal distribution with two variables defined as,

@@

\begin{table}[!htb]
  \begin{center}
    \begin{tabular}{cccccc}
      \hline 
      Particle Type & $\mu_\mathrm{A/Q}$ & $\sigma_\mathrm{A/Q}$ & $\mu_\mathrm{Z}$ &
      $\sigma_\mathrm{Z}$ & $\rho$\\
      \hline\hline 
      ${}^{132}$Sn & 2.64 & 0.0014 & 49.95 & 0.209 & -0.052 \\
    %  $\tin{124}{112}$ & 2.24005 & 0.00150934 & 49.9906 & 0.194804 & -0.0671999 \\
    %  $\tin{112}{124}$ & 2.47995 & 0.00182327 & 49.961 & 0.214745 & -0.0742297 \\
      ${}^{108}$Sn & 2.16 & 0.0015 & 49.99 & 0.207 & -0.059 \\
      \hline
    \end{tabular}
    \caption{Multivariate normal distribution fit parameters for four beams.
    \label{beamParameters}}
  \end{center}
\end{table}
Figure~\ref{fig:beampid132} shows a zoomed in view of the PID centered around the ${}^{132}$Sn beam and Fig.~\ref{fig:beampid108} the ${}^{108}$Sn beam. The red lines represent the cut where particles identified inside the circle represent the beam events which are identified as the good beam events. For both ${}^{132}$Sn an ${}^{108}$Sn a 2.83$\sigma$ cut is taken around the mean values.

@@

Near the edges of the detection volume, the clusters of tracks significantly deviate from the trend of the fitted track as seen in different view of the reconstructed clusters in Fig.~\ref{fig:edge}.  This effect comes from an edge effects where the last pads on the edges of the pad-plane have no neighboring pads containing charge, and therefore the last pad, or in the case of the vertical time bucket spectrum, represents the last known position of the collected charge. In this case the cluster position is biased towards the inside of the TPC. While the number of affected clusters is small as compared with the total number in the track, but the deviation at the end of the track is enough to start causing issues in the momentum reconstruction. Simple cuts were taken to graphically remove these clusters around the left,right, top, and bottom of the TPC. The hits that were cut out satisfied the following conditions,

@@

     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{sideviewHighDenCut.jpg}
         \caption{Side view.}
         \label{fig:sideHigh}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{topviewHighDenCut.jpg}
         \caption{Top view.}
         \label{fig:topHigh}
     \end{subfigure}
        \label{fig:highcut}
        \caption{Views showing the high density and edge cuts.}
        \label{fig:elipsecut}
\end{figure}
The track multiplicity in each event is quite high and since the tracks all originate from a common vertex, the density of tracks near the target region is very high. In this region the track separation is too small for the software to correctly determine and sort the charge information into the appropriate track. The information provided by extra vertex point from external BDC tracking described in \ref{sec:bdc}, provides all the information about the vertex location. The bad quality of hit information near the target region only hurt in the tracking and PID of a track. Hits lying within an semi-ellipsoidal cut around the target are removed from the software and not included in the track and momentum reconstruction. Figure~\ref{fig:elipsecut} shows the extent of the ellipsoidal cut in the high density region, along with the edge cuts as shaded red regions in both views of the TPC.

\section{Beam angle selection}
\label{sec:beamangle}
The incoming secondary beam is deflected by the magnetic field and impinges on the target at some small but significant angle. From the BDCs tracking information, the beam is projected as a straight line right up until entering the magnet.  The beam is then propagated through the magnetic field using a Runge-Kutta integration until reaching the target position. The beam angle on target can be categorized by two angles $\theta_{a_{proj}}$ and $\theta_{b_{proj}}$ defined as,

@@

\begin{equation}
  \theta_\mathrm{a,proj}=\tan^{-1}\frac{p_x}{p_z},\quad
  \theta_\mathrm{b,proj}=\tan^{-1}\frac{p_y}{p_z},
  \label{beamAngle}
\end{equation}

where $p_x$, $p_y$, and $p_z$ are the components of beam momentum vector.

@@

    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{beamAngle-Sn124.png} 
        \caption{Beam angle of the $\tin{124}{112}$ system} \label{fig:beamangle124}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{beamAngle-Sn132.png} 
        \caption{Beam angle of the $\tin{132}{124}$ system} \label{fig:beamangle132}
    \end{subfigure}
\label{fig:beamangle}
\end{figure}

Figure~\ref{fig:beamangle} shows the distribution of beam angles for ${}^{132}$Sn and ${}^{108}$Sn systems. Graphical cuts represented by the red lines where events lying in the cuts represent beams that were reconstructed correctly by the beam tracking software outlined in \cite{jon}. Helping to eliminate some of the poorly reconstructed beam events.

\section{Vertex Cut}

@@

The vertex information of an event, described in \ref{sec:vertex}, is estimated from reconstructing the tracks of an event to one common source. The secondary beam encounters several solid and gaseous materials were a nuclear reaction can occur anywhere along the beam line. Figure~\ref{fig:overviewVertex} shows the z component of the reconstructed vertex for all events in the ${}^{132}$Sn system. Several peaks are seen in the spectrum which correspond to several dense materials in the beam line such as the entrance windows, target frame, Active Veto, with the largest peak representing the target. Also collisions happen with the detector gas inside the TPC volume which are typically called active target events. To ensure that the secondary beam is really on the target of choice we perform a vertex cut around where we believe the vertex location of the target to be.

 The target position was measured to be \SI{-13.2}{\milli\metre}. The z-component of all runs in each beam type are plotted around the the target region in Fig.~\ref{fig:vertexdist}. The mean position of the vertex from the reconstructed data is around \SI{-14.76}{\milli\metre} which is about \SI{1.6}{\milli\metre} off from the expected target position. The mean position for each system is listed in Table~\ref{tb:vertexresol}. Since the target thickness was less than \SI{1}{\milli\metre} for all targets. Neglecting the small distributions in the thin target, the measured width of the vertex distribution can be interpreted directly as the vertex resolution of the TPC. The extracted vertex resolutions of each system are summarized in Table~\ref{tb:vertexresol}, with an average vertex resolution of \SI{1.2}{\centi\metre}. The difference between the measured and actual target location is 10 times smaller than the intrinsic vertex resolution of the detector and is really and insignificant difference.

@@

The impact parameter is a theoretical quantity defined as the distance, between two centers of colliding nuclei. Though there is no way to directly measure the impact parameter in an experiment, the track multiplicity can be indirectly related to the estimated impact parameter of a collision \cite{impactpar}. As discussed in Section \ref{sec:kyoto}, the Kyoto multiplicity array was used to experimentally trigger on central nuclear collision events. This is under the assumption that as the number of charged particles produced in collision is related to the overlap region of the two nuclei. This is best described by a spectator-participant model of nuclear collisions, where a large fraction of the participating nucleons in the overlap region of the two colliding nuclei fragment into individual and clusters of nucleons. In the case where the impact parameter is zero, all the nucleons in both nuclei participate, were as larger impact parameters -- more peripheral collisions -- less nucleons participate.

@@

\section{Track Quality Cuts}
%graphic showing distance to vertex, number of clusters
%number of cluster cut
%distance to vertex
In this section we will discuss cuts which address the quality of the reconstructed track, in the following we will simply refer this as the ``track quality". Assuming tracks are mostly continuous in clusters, i.e. only randomly missing a few clusters, the number of clusters is directly related to the momentum resolution of a track. Tracks with more clusters correspond to better PID and momentum resolution. Upward-going and downward-going tracks, at larger values of $\theta_{Lab}$ are limited by the vertical space of the TPC. In these regions the track length is short and there are few clusters which are reconstructed. This leads to the low efficiency in these regions as discussed earlier in Sec.~\ref{sec:efficiency}. A cut where tracks with the number of clusters $N_{cl} > 20$ are considered quality tracks. Later in Section~\ref{sec:cutvar} the exact choice of the number 20 is discussed in more detail.

@@

For very long tracks, such as low momentum pions, which can make a complete circular path in the TPC, it can happen that the software incorrectly identifies the track as several separate tracks. This can happen due to discontinuities in the track due to missing hits from low energy loss, shadowing due to saturation, or possibly the software algorithm itself. Of course including all of the tracks would be counting the track several times over, and would lead to incorrect particle yields. The software will occasionally associate random disassociated clusters and construct a false track, or a so called ``ghost track". These type of tracks contribute to the background in the PID. In both of the situation of the false track or in the case of a split track the false track origin cannot be traced back to anywhere near the vertex of the event, and the distance-Of-Closest-Approach (dOCA) between the track and vertex is very large.

@@

Since the TPC does not have spherical symmetry, there are geometric constraints and considerations that effect the efficiency of track reconstruction. Most notably upward and downward-going tracks will have the shortest track length and will correspond to low regions of efficiency as seen in \ref{sec:efficiency}. The relevant geometries which determine these regions are the corners of the field cage. Figure~\ref{fig:angleEffExplanation} shows a cartoon drawing of the front of the field cage were the angles of each corner is given. Since the window to the field cage was slightly higher, the upper angles are slightly smaller than the lower angles. These angles are actually the same as the spherical $\phi$ angle, where the left and right, most efficient regions, correspond to  $\ang{0}\substack{+24 \\ -34}$ and $\ang{180}\substack{+34 \\ -24}$.  While the exact area of zero efficiency depends on polar angle $\theta_{Lab}$, a clear cut off around these angles can be seen in Fig.~\ref{fig:angleEffExplanation} in the angular distribution of the measured tracks, where the experimental data was fitted with a Fermi distribution to find the inflection point of where the efficiency drops off significantly. The angular cuts applied to the data is listed by particle type in Table~\ref{tb:anglecuts}. The values are very similar to the simple calculation that was outlined above.

@@

These regions can also be seen in the earlier discussion in the MC embedded efficiency in Fig.~\ref{fig:pim_eff_ex}.  We could of course include a larger region where the efficiency is small but not zero but it is good practice not to efficiency correct already small numbers with small efficiency values. The solid angle covered by the cut is written as,

>>>>

\section{Radio Isotope Beam Factory (RIBF) Facility }
\label{sec:beam}
%Cyclotron facility overview.
%Samurai line overview.
%Beam line element overview.
%Big rips beam PID. reference 
The primary and secondary beams were produced at the Radioactive Isotope Beam Factory (RIFB) facility at RIKEN, in Wako-shi, Japan. The RIBF facility starts with two primary beam types, ${}^{132}$Xe and ${}^{238}$U, which are produced by an ion-source and accelerated to progressively higher kinetic energies by a linear accelerator (RILAC), and 4 different cyclotrons (RRC, fRC, IRC, and SRC), until they reach a beam energy of \SI{345}{\MeVA}. Figure~\ref{fig:samuraiBeamLine} shows the later stages of the cyclotrons and the following beam lines they feed into.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{beamPID.png}
\caption{Overall beam PID for all the systems. Several contaminants other than the desired secondary beam can be seen.}
\label{fig:beampid}
\end{figure}

@@

%Figures of beam contaminants in our beam PID line. 
%Table of beam purity reference to Jon's thesis paper
The secondary beam is produced through the projectile fragmentation of the primary beam off of a \SI{3}{\milli\metre} thick, rotating Be target \cite{inflightsep}. The resulting fragments are filtered in-flight to the desired seconary beam. The in-flight separation is handled by the BigRIPS fragment separator which is shown in Fig.~\ref{fig:samuraiBeamLine}. The dipole magnets D1 and D2 act as a velocity filter, selecting on certain magnetic rigidities $\beta\rho$. Several sets of slits further purify the secondary beam quality by discarding particles which do not focus on the right focal planes. These are the areas where the particles with different velocities focus to different locations in space, which occur at F3,F5, and F7 positions.  Each beam is tracked with the remaining part of the BigRIPS spectrometer tracking system. The particle identification of each beam is achived by the TOF-B$\rho$-$\Delta$E method described in \cite{bigrips}, where the Time of Flight (TOF) information is given by the time it takes to cross two plastic scintilators at F3 and F7 focal planes, and the $\Delta$E information is given by the MUlti-Sampling Ionization Chamber (MUSIC) \cite{music}. From this method the atomic charge, Z, and mass to charge ratio, A/Q, of each particle was measured and separate species represent two-dimensional Gaussian in this space.

Figure~\ref{fig:beampid} shows the beam PID for all the systems for events which satisfied the trigger, several contaminants other than the desired secondary beams still passed through the BigRIPS spectrometer and made it into the TPC. The beam purity of each desired secondary beam is listed in Table~\ref{tb:beams}. The desired secondary beam of interest can be selected by using an appropriate gate around the corresponding particle. Each particle gate is selected by fitting a multivariate normal distribution with two variables defined as,

@@

\begin{table}[!htb]
  \begin{center}
    \begin{tabular}{cccccc}
      \hline 
      Particle Type & $\mu_\mathrm{A/Q}$ & $\sigma_\mathrm{A/Q}$ & $\mu_\mathrm{Z}$ &
      $\sigma_\mathrm{Z}$ & $\rho$\\
      \hline\hline 
      ${}^{132}$Sn & 2.64 & 0.0014 & 49.95 & 0.209 & -0.052 \\
    %  $\tin{124}{112}$ & 2.24005 & 0.00150934 & 49.9906 & 0.194804 & -0.0671999 \\
    %  $\tin{112}{124}$ & 2.47995 & 0.00182327 & 49.961 & 0.214745 & -0.0742297 \\
      ${}^{108}$Sn & 2.16 & 0.0015 & 49.99 & 0.207 & -0.059 \\
      \hline
    \end{tabular}
    \caption{Multivariate normal distribution fit parameters for four beams.
    \label{beamParameters}}
  \end{center}
\end{table}

Figure~\ref{fig:beampid132} shows a zoomed in view of the PID centered around the ${}^{132}$Sn beam and Fig.~\ref{fig:beampid108} the ${}^{108}$Sn beam. The red lines represent the cut where particles identified inside the circle represent the beam events which are identified as the good beam events. For both ${}^{132}$Sn an ${}^{108}$Sn a 2.83$\sigma$ cut is taken around the mean values.

@@

Near the edges of the detection volume, the clusters of tracks significantly deviate from the trend of the fitted track as seen in different view of the reconstructed clusters in Fig.~\ref{fig:edge}.  This happens because the last pads on the edges of the pad-plane have no neighboring pads containing charge, and therefore the last pad represents the last known position of the collected charge. In this case the cluster position is biased towards the inside of the TPC. This also occurs for the first and last time bucket in the vertical direction. While the number of affected clusters is small, compared with the total number in the track, the deviation at the end of the track is enough cause issues in the momentum reconstruction. Simple cuts were taken to graphically remove these clusters around the left,right, top, and bottom of the TPC. The hits that were cut out satisfied the following conditions,

@@

     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{sideviewHighDenCut.jpg}
         \caption{Side view.}
         \label{fig:sideHigh}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{topviewHighDenCut.jpg}
         \caption{Top view.}
         \label{fig:topHigh}
     \end{subfigure}
        \label{fig:highcut}
        \caption{Views showing the high density and edge cuts.}
        \label{fig:elipsecut}
\end{figure}

The track multiplicity in the TPC ranges from 20-80 tracks. Since all the tracks originate from a common vertex, the density of tracks near the target region is very high. In this region the separation between tracks is too small for the software to correctly determine which clusters belong to which tracks, making the information useless or even incorrect. Also, the information provided by extra vertex point from external BDC tracking described in \ref{sec:bdc}, provides all the information about the vertex location. The bad quality of hit information near the target region only hurt in the tracking and PID of a track. Hits lying within an semi-ellipsoidal cut around the target are removed from the software and not included in the track and momentum reconstruction. Figure~\ref{fig:elipsecut} shows the extent of the ellipsoidal cut in the high density region, along with the edge cuts as shaded red regions in both views of the TPC.

\section{Beam angle selection}
\label{sec:beamangle}
The incoming secondary beam is deflected by the magnetic field and impinges on the target at some angle. From the BDC tracking information, the beam is projected as a straight line right up until entering the magnet.  The beam is then propagated through the magnetic field using a Runge-Kutta integration until reaching the target position \cite{jon}. The beam angle on target can be categorized by two angles $\theta_{a_{proj}}$ and $\theta_{b_{proj}}$ where $p_x$, $p_y$, and $p_z$ are the components of beam momentum vector:

@@

\begin{equation}
  \theta_\mathrm{a,proj}=\tan^{-1}\frac{p_x}{p_z},\quad
  \theta_\mathrm{b,proj}=\tan^{-1}\frac{p_y}{p_z}.
  \label{beamAngle}
\end{equation}

@@

    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{beamAngle-Sn124.png} 
        \caption{Beam angle of the $\tin{124}{112}$ system} \label{fig:beamangle124}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{beamAngle-Sn132.png} 
        \caption{Beam angle of the $\tin{132}{124}$ system} \label{fig:beamangle132}
    \end{subfigure}
\caption{Beam angle at the face of the target in the TPC for all 4 systems.}
\label{fig:beamangle}
\end{figure}

Figure~\ref{fig:beamangle} shows the distribution of beam angles for ${}^{132}$Sn and ${}^{108}$Sn systems. Graphical cuts represented by the red lines where events lying in the cuts represent beams that were reconstructed correctly by the beam tracking software outlined in \cite{jon}. This helped to eliminate some of the poorly reconstructed beam events. The angle rotation becomes important in the discussion of transforming back to the center-of-mass system as will be discussed in Section~\ref{sec:pionspectra}.

\section{Vertex Cut}
\label{sec:vertexcut}

@@

The vertex information of an event, described in \ref{sec:vertex}, is estimated from reconstructing the tracks of an event to one common point. The secondary beam encounters several solid and gaseous materials along the beam line where a nuclear reaction can occur. Figure~\ref{fig:overviewVertex} shows the z-component of the reconstructed vertex for all events in the ${}^{132}$Sn system. Several peaks are seen in the spectrum which correspond to several dense materials in the beam line such as the entrance windows, target frame, Active Veto, with the largest peak representing the target. Collisions also happen with the detector gas inside the TPC volume where the gas acts as a target. To ensure that the secondary beam is really on the target of choice we perform a vertex cut around where we believe the vertex location of the target to be.

 The target position was physically expected to be \SI{-13.2}{\milli\metre} in the TPC coordinate frame. The z-component of all runs in each beam type are plotted around the the target region in Fig.~\ref{fig:vertexdist}. The mean position of the vertex from the reconstructed data is around \SI{-14.76}{\milli\metre} which is about \SI{1.6}{\milli\metre} off from the expected target position. The mean position for each system is listed in Table~\ref{tb:vertexresol}. Since the target thickness was less than \SI{1}{\milli\metre} for all targets, we can assume the thickness of the target has a minimal effect on the vertex resolution of the TPC. Therefore the measured width of the vertex distribution can be interpreted directly as the vertex resolution of the TPC. The extracted vertex resolutions of each system are summarized in Table~\ref{tb:vertexresol}, with an average vertex resolution of \SI{1.2}{\centi\metre}. The difference between the measured and actual target location is 10 times smaller than the intrinsic vertex resolution of the detector and is an insignificant difference.

@@

The impact parameter is a theoretical quantity defined as the distance between two centers of colliding nuclei. Though there is no way to directly measure the impact parameter in an experiment, the track multiplicity is indirectly related to the estimated impact parameter of a collision \cite{impactpar}. As discussed in Section \ref{sec:kyoto}, the Kyoto multiplicity array was used to experimentally trigger on central nuclear collision events. This is under the assumption that number of charged particles produced in collision is related to the overlap region of the two nuclei. This is best described by a spectator-participant model of nuclear collisions, where a large fraction of the participating nucleons in the overlap region of the two colliding nuclei fragment into individual and clusters of nucleons. In the case where the impact parameter is zero, all the nucleons in both nuclei participate, were as larger impact parameters -- more peripheral collisions -- less nucleons participate.

@@

\section{Track Quality Cuts}
\label{sec:qualitycut}
%graphic showing distance to vertex, number of clusters
%number of cluster cut
%distance to vertex
In this section we will discuss cuts which address the quality of the reconstructed track, in the following we will simply refer this as the \emph{track quality}. Assuming tracks are mostly continuous in clusters, i.e. only randomly missing a few clusters, the number of clusters is directly related to the momentum resolution of a track. Tracks with more clusters correspond to better PID and momentum resolution. Upward-going and downward-going tracks, at larger values of $\theta_{Lab}$ are limited by the vertical space of the TPC. In these regions the track length is short and there are few clusters which are reconstructed. This leads to the low efficiency in these regions as discussed earlier in Sec.~\ref{sec:efficiency}. A cut where tracks with the number of clusters $N_{cl} > 20$ are considered quality tracks. Later in Section~\ref{sec:cutvar} the exact choice of the number 20 is discussed in more detail.

@@

For very long tracks, such as low momentum pions, which can make a complete circular path in the TPC, it can happen that the software incorrectly identifies the track as several separate tracks. This can happen due to discontinuities in the track due to missing hits from low energy loss, shadowing due to saturation, or possibly the software algorithm itself. Including all of the tracks would be counting the track several times over, and would lead to incorrect particle yields. The software can also occasionally associate random disassociated clusters and construct a false track, or a so called \emph{ghost track}. These type of tracks contribute to the background in the PID. In both of these situations the track origin cannot be traced back to anywhere near the vertex of the event, and the distance-Of-Closest-Approach (dOCA) between the track and vertex is very large.

@@

Since the TPC does not have spherical symmetry, there are regions of bad geometric acceptance that effect the efficiency of track reconstruction. Most notably upward and downward-going tracks will have the shortest track length and will correspond to low regions of efficiency as seen in \ref{sec:efficiency}. The relevant geometries which determine these regions are the corners of the field cage. Figure~\ref{fig:angleEffExplanation} shows a cartoon drawing of the front of the field cage were the angles of each corner is given. Since the window to the field cage was slightly higher, the upper angles are smaller than the lower angles. These angles are actually the same as the spherical $\phi$ angle, where the left and right, are most efficient regions, correspond to  $\ang{0}\substack{+24 \\ -34}$ and $\ang{180}\substack{+34 \\ -24}$.  While the exact area of zero efficiency depends on polar angle $\theta_{Lab}$, a clear cut off around these angles can be seen in Fig.~\ref{fig:angleEffExplanation} in the angular distribution of the measured tracks. Here, the experimental data was fitted with a Fermi distribution to find the inflection point of where the efficiency drops off significantly. The angular cuts applied to the data is listed by particle type in Table~\ref{tb:anglecuts}. The values are very close to the simple calculation outlined previously.

@@

These regions can also be seen in the earlier discussion in the MC embedded efficiency in Fig.~\ref{fig:pim_eff_ex}.  We could of course include a larger region where the efficiency is small, but not zero, but it is not good practice to  correct using small efficiency values. Instead we apply a reasonable $\phi$ cut in the good areas of reasonable efficiency and correct for the regions excluded from the cut. The solid angle covered by the cut is written as,

### 2020-07-04 chapters/ch_results.tex
Second draft. Intro needs work

To correct for the efficiency, we apply a track-by-track correction, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the database, from the parameters discussed in Section~\ref{sec:efficiency}. The correction factor $C_i$ is then defined as,

@@

Each track that is identified as a pion as describe in Section~\ref{sec:pid}, is then weighted by the correction factor $C_i$, when filling the histogram of any observable. In this way, we can correct for the efficiency of the track into any observable of interest, notably transforming from the lab frame into the center-of-mass (CM) frame. Before performing the transformation to the CM system, recall there is a small, but non-negligible, beam angle in the Lab frame; see Section~\ref{sec:beamangle}. The beam angle for each event is measured by the beam tracking software and a rotation is applied to all the tracks in an event, to align the beam angle along the z-axis. Doing so makes the transformation into the CM system much simpler to describe. If the beam direction is defined by a unit vector $\hat{b}$, we can define the rotation that rotates the beam into the z-axis as a rotation about an arbitrary vector $\hat{v} = \hat{b}\times\hat{z}$ where the angle between the two is given by $\cos \theta = \hat{b}\cdot\hat{z}$. A rotation is then applied around the vector $\hat{v}$.

Once all the events have been rotated to align with the z-axis, transforming from the Lab to the CM frame is done by a Lorentz transformation. The 4-momentum vector in the lab frame is defined as $\textbf{P} = (E/c,p_x,p_y,p_z)$. Where the corresponding Lorentz transform into the CM frame along the beam (z-axis) is defined as,

@@

where $\beta$, describes the velocity of the CM system, and $\gamma=\sqrt{1-\beta^2}^{-1}$. The parameter $\beta$ can be determined from the total momentum of the system in the Laboratory frame $P = \sqrt{ T_{P}^2 - M_{T}^2}$ and the total energy of the system $E = T_{P} + M_{P} + M_{T}$ as $\beta = -P/E$, where the (-) sign denotes the correct direction for transforming from the Lab to the CM frame. The CM transformed track is defined as $p^{CM} = \textbf{A}p^{Lab}$.

@@

Figure~\ref{fig:pionspectra} shows pion CM kinetic energy spectra for both the $\tin{132}{124}$ and $\tin{108}{112}$ systems; corrected for efficiency and accounting for the solid angle of 4$\pi$. This data marks the first time pion spectral data  has been measured at sub-threshold energies. In the pion spectra, we can see the effect of the coulomb potential, which accelerates $\pi^+$ and decelerate $\pi^-$ particles, due to the positive charge of the nuclear medium. The low $\pi^+$ production at low energy is caused by the coulomb barrier between protons. The effects of the Coulomb force are the largest effects on the spectra of the pions and will play an important role when looking at pion spectral ratios.

@@

The data taken has been compared with 7 commonly used transport theoretical codes. These codes took part in a large collaborative effort in order to standarize certain common elements of each codes. Each comparison project simulating nuclear matter in a box type simulation with fixed, well defined, initial conditions. Though not physical, this provided a benchmark for codes to compare to; the solutions of the total number of pions and $\Delta$ produced had an analytical solution one would expect from statistical equalibrium, providing a good benchmark. This allowed for each code to systematically go through the numerical treatment and details in each code such as intialization of nuclei, stability of the code, numerical handing of pauli-blocking, etc. \cite{theoryComp1,theoryComp2}. These codes were taken in their best configuration, as a result of these comparison projects, without any prior knowledge of the experimental data. We then simulated the 4 systems measured in the \spirit TPC at an impact parameter of \SI{3}{\femto\metre} at \SI{270}{\MeVA} beam energy. While numerical treatments of each code are resonably similar, each code differs in its treatment of pion and $\Delta$ dynamics. Some codes contain modifications to how the pion behaves in nuclear matter, i.e. \emph{in-medium effects}, typically introduced by including a pion optical potential which describes the pion scattering and absorptions. Some codes include the iso-scalar and the iso-vector delta potential, which are not very well constrained but are important in the production of pions CITE HERE. The set of 7 codes represent the best codes which can simulate pions at these energies. As will be the re-occuring theme, there is a large variation between theoretical codes; greater than the variation between different symmetry energy in a particular code. Though many of the numerical and theoretical treatments have been addressed, many of the details of pion production and dynamics have not been addressed. This is an extremely complex task requiring major theoretical efforts and new data. We expect the data in this thesis and in \cite{jon} to provide much needed experimental data in which each code may compare too. The total theoretical uncertainty is too large relative to the experimental error bars to make a constraint on the density dependence of the symmetry energy. Any constraint would depend on the particular code one is using and would result in conflicting conclusions when using another theory. There is no good consensus at this time which codes should be favored over other codes, as the magnitude of the effects included in each code are still debated in order of importance. Here I will present the current level of agreement with the codes and remark on some of the implications it has moving forward.

>>>>

To correct for the efficiency, we apply a track-by-track correction, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the efficiency database, from the parameters discussed in Section~\ref{sec:efficiency}. The correction factor $C_i$ is defined as,

@@

Each track that is identified as a pion as described in Section~\ref{sec:pid}, is then weighted by the correction factor $C_i$, when filling the histogram of any observable. In this way, we can correct for the efficiency on a track-by-track basis, allowing for any transformation of the track of interest into any observable, notably transforming from the lab frame into the center-of-mass (CM) frame. Before performing the transformation to the CM system, recall there is a small, but non-negligible, beam angle in the Lab frame discussed in Section~\ref{sec:beamangle}. First we rotate each event so that the beam will align with the z-axis. Doing so makes the transformation into the CM system much simpler to describe. If the beam direction is defined by a unit vector $\hat{b}$, we can define the rotation that rotates the beam into the z-axis as a rotation about an arbitrary vector $\hat{v} = \hat{b}\times\hat{z}$ where the angle between the two is given by $\cos \theta = \hat{b}\cdot\hat{z}$. A rotation with the angle $\theta$ is then applied around the vector $\hat{v}$.

Once all the events have been rotated to align with the z-axis, transforming from the Lab to the CM frame is done by a Lorentz transformation. The 4-momentum vector in the lab frame is defined as $\textbf{P} = (E/c,p_x,p_y,p_z)$. Where the corresponding Lorentz transform into the CM frame along the beam (z-axis) is defined as the Lorentz transformation matrix $A$,

@@

where $\beta$, describes the velocity of the CM system, and $\gamma=\sqrt{1-\beta^2}^{-1}$. The parameter $\beta$ can be determined from the total momentum of the system in the Laboratory frame $P = \sqrt{ T_{P}^2 + 2M_{T}T_{T}}$ and the total energy of the system $E = T_{P} + M_{P} + M_{T}$ -- where $T_{P}$ is the projectile kinetic energy in the Lab frame, $M_{T}$, and $M_{P}$ are the total mass of the target and projectile, where $\beta = -P/E$; the (-) sign denotes the correct direction for transforming from the Lab to the CM frame. The transformation of each  track is defined as $p^{CM} = \textbf{A}p^{Lab}$.

@@

Figure~\ref{fig:pionspectra} shows pion CM kinetic energy spectra for both the $\tin{132}{124}$ and $\tin{108}{112}$ systems; corrected for efficiency and accounting for the solid angle of 4$\pi$. This data marks the first time pion spectral data  has been measured at sub-threshold energies. In the pion spectra, we can see the effect of the coulomb potential, which accelerates $\pi^+$ and decelerate $\pi^-$ particles, due to the positive charge of the nuclear medium. The low $\pi^+$ production at low energy is caused by the coulomb barrier between protons which limits $\pi^+$ production at low energies. The Coulomb force has the largest effect on the spectra of the pions and will play an important role when looking at pion spectral ratios. It has almost no role in the total pion yield.

@@

The data taken has been compared with 7 commonly used transport theoretical codes. These codes took part in a large collaborative effort in order to standarize certain common elements of each codes. Each comparison project simulated nuclear matter in a box type simulation with fixed, well defined, initial conditions. Though not physical, this provided a benchmark for codes to compare to; the solutions of the total number of pions and $\Delta$ produced had an analytical solution one would expect from statistical equalibrium, providing a good benchmark. This allowed for each code to systematically go through the numerical treatment and details in each code such as intialization of nuclei, stability of the code, numerical handing of pauli-blocking, etc. \cite{theoryComp1,theoryComp2}. These codes were taken in their best configuration, as a result of these comparison projects, without any prior knowledge of the experimental data. We then simulated the 4 systems measured in the \spirit TPC at an impact parameter of \SI{3}{\femto\metre} at \SI{270}{\MeVA} beam energy. While the numerical treatments of each code are resonably similar, each code differs considerably in their treatment of pion and $\Delta$ dynamics. Some codes contain modifications to how the pion behaves in nuclear matter, i.e. \emph{in-medium effects}, typically introduced by including a pion optical potential which describes the pion scattering and absorptions. Some codes include the iso-scalar and the iso-vector delta potential, which are not very well constrained but are important in the production of pions CITE HERE. The set of 7 codes represent the best codes which can simulate pions at these energies.

As will be the re-occuring theme, there is a large variation between theoretical codes; greater than the variation between different symmetry energy in a particular code. Though many of the numerical and theoretical treatments have been addressed, many of the details of pion production and dynamics have not been addressed. This is an extremely complex task requiring major theoretical efforts and new data. We expect the data in this thesis and in \cite{jon} to provide much needed experimental data in which each code may compare too. The total theoretical uncertainty is too large relative to the experimental error bars to make a constraint on the density dependence of the symmetry energy. Any constraint would depend on the particular code one is using and would result in conflicting conclusions when using another theory. There is no good consensus at this time which codes should be favored over other codes, as the magnitude of the effects included in each code are still debated in order of importance. Here I will present the current level of agreement with the codes and remark on some of the implications it has moving forward.
